---
title: "add DESeq2 to mass spec data"
author: "Karen Chu"
date: "12/19/2019"
output: html_document
---

LSC MSI2 KO DESeq2 data append to LSC mass spec data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries.

```{r libraries}
library(openxlsx)
library(DESeq2)
library(dplyr)
library(tidyr)
```

Set work dir

```{r workdir}
folder <- "~/mount/chuk/LSC_LSK_mass_spec/"
output.folder <- "~/mount/chuk/LSC_LSK_mass_spec/output/"
plots.folder <- "~/mount/chuk/LSC_LSK_mass_spec/figures/"
```

Import mass spec data.

```{r data}
setwd(output.folder)

lsc.intensity <- readRDS("lsc.intensity_significance.rds")
lsc.normal.intensity <- readRDS("lsc.normal.intensity_significance.rds")
lsc.lfq <- readRDS("lsc.lfq_significance.rds")

# Add gene name column
lsc.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsc.intensity) )
lsc.normal.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsc.normal.intensity) )
lsc.lfq$gene.symbol <- gsub( "\\...*", "", rownames(lsc.lfq) )
```

For rows with multiple gene names, separate gene names and duplicate row data for them.

Old method for separating multiple gene rows:
s <- strsplit(lsc.intensity.multiple.gene.row$gene.symbol, split = ";")
s1 <- data.frame(V1 = rep(lsc.intensity.multiple.gene.row$gene.symbol, sapply(s, length)), 
                 V2 = unlist(s))

```{r multiple gene names rows}
split.multiple.gene.rows <- function(df) {
  
  # Identify rows with multiple genes
  multiple.gene.row.index <- which(grepl(";", df$gene.symbol))
  df.multiple.gene.row <- df[ multiple.gene.row.index, ]
  df.remove.multiple.gene.row <- df[ -multiple.gene.row.index, ]

  # Split multiple genes into their own rows and duplicate data.
  split.multiple.genes <- df.multiple.gene.row %>% separate_rows(gene.symbol)

  # Re-bind multiple gene rows back to original data.
  final.data <- rbind(df.remove.multiple.gene.row, split.multiple.genes)
  
  return(final.data)
  
}

lsc.intensity <- split.multiple.gene.rows(lsc.intensity)
lsc.normal.intensity <- split.multiple.gene.rows(lsc.normal.intensity)
lsc.lfq <- split.multiple.gene.rows(lsc.lfq)
```

Import rna-seq MSI2 KO LSC data.

```{r import hypertribe}
setwd("~/mount/chuk/data_resources/")

msi2.ko.lsc <- read.xlsx("RNA-seq_HSC_LSC_GMP_MSI2_021214.xlsx", sheet = 2)
```

Import rna-seq and hypertribe data.

```{r import hypertribe}
setwd("~/mount/chuk/data_resources/")

lsc.hypertribe <- read.xlsx("mouse_lsc_snp_counts_dedupped_significant.xlsx")

lsc.hypertribe.subset <- lsc.hypertribe %>% select(c("chr", "pos", "strand", 
                                                     "gene.symbol", "diff.frequency", "ADA.frequency",
                                                     "DCD.frequency", "MIG.frequency", "gene.num.edits",
                                                     "ADA.fpkm", "DCD.fpkm", "MIG.fpkm"))
```

Change DESeq2 column names to get Gene column as "gene.symbol"

```{r gene.symbol}
colnames(msi2.ko.lsc)[1] <- "gene.symbol"
```

Merge DESeq2 data with mass spec data.

```{r merge}
# Merge data
msi2.ko.lsc.subset <- msi2.ko.lsc [ msi2.ko.lsc$gene.symbol %in% unique(lsc.intensity$gene.symbol), ]
lsc.intensity.deseq <- merge(lsc.intensity, msi2.ko.lsc.subset, by="gene.symbol", all=TRUE)

msi2.ko.lsc.subset <- msi2.ko.lsc [ msi2.ko.lsc$gene.symbol %in% unique(lsc.normal.intensity$gene.symbol), ]
lsc.normal.intensity.deseq <- merge(lsc.normal.intensity, msi2.ko.lsc.subset, by="gene.symbol", all=TRUE)

msi2.ko.lsc.subset <- msi2.ko.lsc [ msi2.ko.lsc$gene.symbol %in% unique(lsc.lfq$gene.symbol), ]
lsc.lfq.deseq <- merge(lsc.lfq, msi2.ko.lsc.subset, by="gene.symbol", all=TRUE)

# Rename column names
colnames(lsc.intensity.deseq)[18:25] <- paste("DESeq2", colnames(lsc.intensity.deseq[18:25]), sep=".")
colnames(lsc.normal.intensity.deseq)[18:25] <- paste("DESeq2", colnames(lsc.normal.intensity.deseq[18:25]), sep=".")
colnames(lsc.lfq.deseq)[18:25] <- paste("DESeq2", colnames(lsc.lfq.deseq[18:25]), sep=".")
```

Merge with hypertribe data.

```{r append rnaseq}
append.hypertribe.data <- function(df.mass.spec, df.hypertribe) {
  
  # Subset only genes that appear in mass spec
  # If gene in mass spec but not in hypertribe, then it will get NA for hypertribe data
  shared.genes <- intersect(as.character(df.mass.spec$gene.symbol), 
                            as.character(df.hypertribe$gene.symbol))
  df.hypertribe.subset <- df.hypertribe [ df.hypertribe$gene.symbol %in% shared.genes, ]
  
  df.final <- merge( df.mass.spec, df.hypertribe.subset, by="gene.symbol", all=TRUE )
  
  return(df.final)
  
}

lsc.intensity.deseq.hypertribe <- append.hypertribe.data(lsc.intensity.deseq, lsc.hypertribe.subset)
lsc.normal.intensity.deseq.hypertribe <- append.hypertribe.data(lsc.normal.intensity.deseq, lsc.hypertribe.subset)
lsc.lfq.deseq.hypertribe <- append.hypertribe.data(lsc.lfq.deseq, lsc.hypertribe.subset)
```

Write output files.

```{r write}
setwd(output.folder)

# Write
write.csv(lsc.intensity.deseq.hypertribe, "lsc.intensity_with.LSC.msi2.ko.deseq2.and.hypertribe.data.csv", row.names = FALSE)
write.csv(lsc.normal.intensity.deseq.hypertribe, "lsc.normal.intensity_with.LSC.msi2.ko.deseq2.and.hypertribe.data.csv", row.names = FALSE)
write.csv(lsc.lfq.deseq.hypertribe, "lsc.lfq_with.LSC.msi2.ko.deseq2.and.hypertribe.data.csv", row.names = FALSE)
```










