---
title: "LSC LSK mass spec"
author: "Karen Chu"
date: "12/13/2019"
output: html_document
---

Hypothesis: MSI2 regulates translation of its targets.

Methods: Perform mass spec on proteins present in MSI2 WT vs MSI2 KO.

Change log2fc to +1 to avoid infinity
Try to decrease the number of tests to improve p-adjusted by putting a threshold on difference between cre pos +tam and cre pos -tam.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries.

```{r libraries}
library(openxlsx)
library(ggplot2)
```

Set work dir

```{r work dir}
folder <- "~/mount/chuk/LSC_LSK_mass_spec/"
output.folder <- "~/mount/chuk/LSC_LSK_mass_spec/output/"
plots.folder <- "~/mount/chuk/LSC_LSK_mass_spec/figures/"
```

Import data.

```{r data}
setwd(folder)

lsc <- read.xlsx("data/Msi2KOLSC_MS.xlsx")
lsk <- read.xlsx("data/Msi2KO_3wks_LSK_MS.xlsx")

# Remove NA (blank gene name entries in xlsx)
lsc <- lsc [ !(is.na(lsc$Gene.names)), ]
lsk <- lsk [ !(is.na(lsk$Gene.names)), ]

rownames(lsc) <- paste( lsc$Gene.names, seq(1, nrow(lsc)), sep="..." )
rownames(lsk) <- paste( lsk$Gene.names, seq(1, nrow(lsk)), sep="..." )
```

Subset data into the different intensity data (for better organization)

```{r separate}
# Intensity
lsc.intensity.index <- grep("^Intensity", colnames(lsc))
lsk.intensity.index <- grep("^Intensity", colnames(lsk))
lsc.intensity <- lsc [ ,lsc.intensity.index ]
lsk.intensity <- lsk [ ,lsk.intensity.index ]

# Normal intensity
lsc.normal.intensity.index <- grep("^norm_Intensity", colnames(lsc))
lsk.normal.intensity.index <- grep("^norm_Intensity", colnames(lsk))
lsc.normal.intensity <- lsc [ ,lsc.normal.intensity.index ]
lsk.normal.intensity <- lsk [ ,lsk.normal.intensity.index ]

# LFQ
lsc.lfq.index <- grep("^LFQ", colnames(lsc))
lsk.lfq.index <- grep("^LFQ", colnames(lsk))
lsc.lfq <- lsc [ ,lsc.lfq.index ]
lsk.lfq <- lsk [ ,lsk.lfq.index ]
```

Calculate mean of all replicates per gene.
Note: You need to escape special characters twice, once for R and once for the regular expression.

```{r mean}
# Calculate log2FC
calculate.log2fc.lsc <- function(df) {
  
  df.cre.pos.plus.tam <- df[ ,grep("y_TAM_cre\\+", colnames(df)) ]
  df.cre.pos.no.tam <- df[ ,grep("noTAM_cre\\+", colnames(df)) ]

  df.cre.neg.plus.tam <- df[ ,grep("y_TAM_cre-", colnames(df)) ]
  df.cre.neg.no.tam <- df[ ,grep("Intensity_noTAM.*re-", colnames(df)) ]
  
  # Calculate mean across replicates
  df.cre.pos.plus.tam.rowMeans <- rowMeans(df.cre.pos.plus.tam, na.rm=TRUE)
  df.cre.pos.no.tam.rowMeans <- rowMeans(df.cre.pos.no.tam, na.rm=TRUE)
  df.cre.neg.plus.tam.rowMeans <- rowMeans(df.cre.neg.plus.tam, na.rm=TRUE)
  df.cre.neg.no.tam.rowMeans <- rowMeans(df.cre.neg.no.tam, na.rm=TRUE)
  
  # Calculate log2FC
  df.cre.pos.log2fc <- log2( df.cre.pos.plus.tam.rowMeans / df.cre.pos.no.tam.rowMeans )
  df.cre.neg.log2fc <- log2( df.cre.neg.plus.tam.rowMeans / df.cre.neg.no.tam.rowMeans )
  
  res <- data.frame( cre.pos.log2fc = df.cre.pos.log2fc,
                     cre.neg.log2fc = df.cre.neg.log2fc )
  
  return(res)
  
}

calculate.log2fc.lsk <- function(df) {
  
  df.cre.pos <- df[ ,grep("re\\+", colnames(df)) ]
  df.cre.neg <- df[ ,grep("re\\-", colnames(df)) ]
  
  # Calculate mean across replicates
  df.cre.pos.rowMeans <- rowMeans(df.cre.pos, na.rm=TRUE)
  df.cre.neg.rowMeans <- rowMeans(df.cre.neg, na.rm=TRUE)
  
  # Calculate log2FC
  df.log2fc <- log2( df.cre.pos.rowMeans / df.cre.neg.rowMeans )

  res <- data.frame( log2fc = df.log2fc )
  
  return(res)
  
}

lsc.intensity.log2fc <- calculate.log2fc.lsc(lsc.intensity)
lsc.normal.intensity.log2fc <- calculate.log2fc.lsc(lsc.normal.intensity)
lsc.lfq.log2fc <- calculate.log2fc.lsc(lsc.lfq)

lsk.intensity.log2fc <- calculate.log2fc.lsk(lsk.intensity)
lsk.normal.intensity.log2fc <- calculate.log2fc.lsk(lsk.normal.intensity)
lsk.lfq.log2fc <- calculate.log2fc.lsk(lsk.lfq)
```

p-value calculated by taking the raw intensity numbers in +tam vs -tam.

Old code:
keepers.index <- which( (abs( rowMeans( df[,cre.pos.tam.index] ) - rowMeans( df[,cre.pos.no.tam.index] ) ) > min.value) )
keepers.index <- which( (abs( rowMeans( df[,cre.neg.tam.index] ) - rowMeans( df[,cre.neg.no.tam.index] ) ) > min.value) )

```{r p-value calculate}
setwd(output.folder)

# Calculate p-value
calculate.pvalue.lsc <- function(df) {
  
  cre.pos.tam.index <- grep("y_TAM_cre\\+", colnames(df))
  cre.pos.no.tam.index <- grep("noTAM_cre\\+", colnames(df))
  
  cre.neg.tam.index <- grep("y_TAM_cre\\-", colnames(df))
  cre.neg.no.tam.index <- grep("noTAM.*re\\-", colnames(df))

  cre.pos.pvalue <- sapply(1:nrow(df), function(i) t.test(df[ i, cre.pos.tam.index ], df[ i, cre.pos.no.tam.index])$p.value)
  cre.neg.pvalue <- sapply(1:nrow(df), function(i) t.test(df[ i, cre.neg.tam.index ], df[ i, cre.neg.no.tam.index])$p.value)

  df.stats <- data.frame( cre.pos.pvalue = cre.pos.pvalue,
                          cre.neg.pvalue = cre.neg.pvalue )
  
  return(df.stats)
  
}

calculate.pvalue.lsk <- function(df) {
  
  cre.pos.index <- grep("cre\\+", colnames(df))
  cre.neg.index <- grep("*re\\-", colnames(df))

  pvalue <- sapply(1:nrow(df), function(i) t.test(df[ i, cre.pos.index ], df[ i, cre.neg.index])$p.value)

  df.stats <- data.frame( pvalue = pvalue )
  
  return(df.stats)
  
}

lsc.intensity.stats <- cbind( lsc.intensity, lsc.intensity.log2fc, calculate.pvalue.lsc(lsc.intensity) )
lsc.normal.intensity.stats <- cbind( lsc.normal.intensity, lsc.normal.intensity.log2fc, calculate.pvalue.lsc(lsc.normal.intensity) )
lsc.lfq.stats <- cbind( lsc.lfq, lsc.lfq.log2fc, calculate.pvalue.lsc(lsc.lfq) )

lsk.intensity.stats <- cbind( lsk.intensity, lsk.intensity.log2fc, calculate.pvalue.lsk(lsk.intensity) )
lsk.normal.intensity.stats <- cbind( lsk.normal.intensity, lsk.normal.intensity.log2fc, calculate.pvalue.lsk(lsk.normal.intensity) )
lsk.lfq.stats <- cbind( lsk.lfq, lsk.lfq.log2fc, calculate.pvalue.lsk(lsk.lfq) )
```

Calculate p-adjusted

```{r p.adjust}
# Calculate p-adjusted
calculate.padjust.lsc <- function(df) {
  
  # Select intensity rows with rowMean >= [set value]
  cre.pos.tam.index <- grep("y_TAM_cre\\+", colnames(df))
  cre.pos.no.tam.index <- grep("noTAM_cre\\+", colnames(df))
  
  # Obtain index of all proteins that have 2 or more replicates with value of 0
  # Calculate number of zeros in each row
  keepers.index.cre.pos.tam <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.pos.tam.index ] == 0 ) ) ) )
  
  # Identify which row has less than 2 zeros
  keepers.index.cre.pos.tam <- which( keepers.index.cre.pos.tam < 2 )
  
  # Repeat with no tam samples.
  keepers.index.cre.pos.no.tam <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.pos.no.tam.index ] == 0 ) ) ) )
  keepers.index.cre.pos.no.tam <- which( keepers.index.cre.pos.no.tam < 2 )
  
  # Get samples that have at least 2 replicates that are not equal to zero in each condition.
  keepers.index <- intersect( keepers.index.cre.pos.tam, keepers.index.cre.pos.no.tam )
  
  # Calculate p-adjusted on specified samples above.
  df$cre.pos.p.adjust <- NA
  df$cre.pos.p.adjust[keepers.index] <- p.adjust(df$cre.pos.pvalue[keepers.index], 'BH')
  
  # Repeat above method with cre negative samples.
  cre.neg.tam.index <- grep("y_TAM_cre\\-", colnames(df))
  cre.neg.no.tam.index <- grep("noTAM.*re\\-", colnames(df))
  
  keepers.index.cre.neg.tam <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.neg.tam.index ] == 0 ) ) ) )
  keepers.index.cre.neg.tam <- which( keepers.index.cre.neg.tam < 2 )
  keepers.index.cre.neg.no.tam <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.neg.no.tam.index ] == 0 ) ) ) )
  keepers.index.cre.neg.no.tam <- which( keepers.index.cre.neg.no.tam < 2 )
  keepers.index <- intersect( keepers.index.cre.neg.tam, keepers.index.cre.neg.no.tam )
  
  df$cre.neg.p.adjust <- NA
  df$cre.neg.p.adjust[keepers.index] <- p.adjust( df$cre.neg.pvalue[keepers.index], 'BH' )
  
  return(df)
  
}

calculate.padjust.lsk <- function(df) {
  
  cre.pos.index <- which( grepl("re\\+", colnames(df)))
  cre.neg.index <- which( grepl("re\\-", colnames(df)))
  
  keepers.index.cre.pos <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.pos.index ] == 0 ) ) ) )
  keepers.index.cre.pos <- which( keepers.index.cre.pos < 2 )
  keepers.index.cre.neg <- unlist( sapply(1:nrow(df), function(i) length( which( df[i, cre.neg.index ] == 0 ) ) ) )
  keepers.index.cre.neg <- which( keepers.index.cre.neg < 2 )
  keepers.index <- intersect( keepers.index.cre.pos, keepers.index.cre.neg )  
  
  df$p.adjust <- NA
  df$p.adjust[keepers.index] <- p.adjust( df$pvalue[keepers.index], 'BH' )
  
  return(df)
  
}

lsc.intensity <- calculate.padjust.lsc(lsc.intensity.stats)
lsc.normal.intensity <- calculate.padjust.lsc(lsc.normal.intensity.stats)
lsc.lfq <- calculate.padjust.lsc(lsc.lfq.stats)

lsk.intensity <- calculate.padjust.lsk(lsk.intensity.stats)
lsk.normal.intensity <- calculate.padjust.lsk(lsk.normal.intensity.stats)
lsk.lfq <- calculate.padjust.lsk(lsk.lfq.stats)
```

Save output.

```{r write}
setwd(output.folder)

write.csv(cbind( lsc$Gene.names, lsc.intensity), "lsc.intensity_significance.csv", row.names = FALSE)
write.csv(cbind( lsc$Gene.names, lsc.normal.intensity), "lsc.normal.intensity_significance.csv", row.names = FALSE)
write.csv(cbind( lsc$Gene.names, lsc.lfq), "lsc.lfq_significance.csv", row.names = FALSE)

write.csv(cbind( lsk$Gene.names, lsk.intensity), "lsk.intensity_significance.csv", row.names = FALSE)
write.csv(cbind( lsk$Gene.names, lsk.normal.intensity), "lsk.normal.intensity_significance.csv", row.names = FALSE)
write.csv(cbind( lsk$Gene.names, lsk.lfq), "lsk.lfq_significance.csv", row.names = FALSE)
```

```{r rds}
setwd(output.folder)

saveRDS(lsc.intensity, "lsc.intensity_significance.rds")
saveRDS(lsc.normal.intensity, "lsc.normal.intensity_significance.rds")
saveRDS(lsc.lfq, "lsc.lfq_significance.rds")

saveRDS(lsk.intensity, "lsk.intensity_significance.rds")
saveRDS(lsk.normal.intensity, "lsk.normal.intensity_significance.rds")
saveRDS(lsk.lfq, "lsk.lfq_significance.rds")
```

Import mass spec data.

```{r data}
setwd(output.folder)

lsc.intensity <- readRDS("lsc.intensity_significance.rds")
lsc.normal.intensity <- readRDS("lsc.normal.intensity_significance.rds")
lsc.lfq <- readRDS("lsc.lfq_significance.rds")

lsk.intensity <- readRDS("lsk.intensity_significance.rds")
lsk.normal.intensity <- readRDS("lsk.normal.intensity_significance.rds")
lsk.lfq <- readRDS("lsk.lfq_significance.rds")

# Add gene name column
lsc.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsc.intensity) )
lsc.normal.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsc.normal.intensity) )
lsc.lfq$gene.symbol <- gsub( "\\...*", "", rownames(lsc.lfq) )

lsk.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsk.intensity) )
lsk.normal.intensity$gene.symbol <- gsub( "\\...*", "", rownames(lsk.normal.intensity) )
lsk.lfq$gene.symbol <- gsub( "\\...*", "", rownames(lsk.lfq) )
```

Histogram log2FC

```{r histogram}
setwd(plots.folder)

histogram.plot.lsc <- function(df, title) {
  
  df.subset <- df [ !(is.na(df$cre.pos.pvalue)), ]
  df.subset <- df.subset [ !(is.na(df.subset$cre.pos.log2fc)), ]
  df.subset <- df.subset [ df.subset$cre.pos.log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$cre.pos.log2fc != "-Inf", ]
  p.pos <- ggplot(df.subset, aes(x=cre.pos.log2fc)) + 
  geom_histogram(color="white", fill="black") +
  ggtitle(paste0(title, " Cre+ log2FC")) +
  theme_minimal() +
  xlim(-10,10) +
  xlab("\nlog2(TAM/noTAM)") +
  ylab("Count\n") +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( "Histogram_", title, "_log2FC_cre.pos.png"), 1000, 800)
  print(p.pos)
  dev.off()
  
  df.subset <- df [ !(is.na(df$cre.neg.pvalue)), ]
  df.subset <- df.subset [ !(is.na(df.subset$cre.neg.log2fc)), ]
  df.subset <- df.subset [ df.subset$cre.neg.log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$cre.neg.log2fc != "-Inf", ]
  p.neg <- ggplot(df.subset, aes(x=cre.neg.log2fc)) + 
  geom_histogram(color="white", fill="black") +
  ggtitle(paste0(title, " Cre- log2FC")) +
  theme_minimal() +
  xlim(-10,10) +
  xlab("\nlog2(TAM/noTAM)") +
  ylab("Count\n") +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( "Histogram_", title, "_log2FC_cre.neg.png"), 1000, 800)
  print(p.neg)
  dev.off()
  
}

histogram.plot.lsk <- function(df, title) {
  
  df.subset <- df [ !is.na(df$pvalue), ]
  df.subset <- df.subset [ !(is.na(df.subset$log2fc)), ]
  df.subset <- df.subset [ df.subset$log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$log2fc != "-Inf", ]
  p <- ggplot(df.subset, aes(x=log2fc)) + 
  geom_histogram(color="white", fill="black") +
  ggtitle(paste0(title, " log2FC")) +
  theme_minimal() +
  xlim(-10,10) +
  xlab("\nlog2(Cre+/Cre-)") +
  ylab("Count\n") +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( "Histogram_", title, "_log2FC.png"), 1000, 800)
  print(p)
  dev.off()

}

histogram.plot.lsc(lsc.intensity, "lsc.intensity")
histogram.plot.lsc(lsc.normal.intensity, "lsc.normal.intensity")
histogram.plot.lsc(lsc.lfq, "lsc.lfq")

histogram.plot.lsk(lsk.intensity, "lsk.intensity")
histogram.plot.lsk(lsk.normal.intensity, "lsk.normal.intensity")
histogram.plot.lsk(lsk.lfq, "lsk.lfq")
```

Volcano plots.

```{r volcano plots}
setwd(plots.folder)

volcano.plot.lsc <- function(df, title, p.value.threshold) {
  
  volcano.input <- as.data.frame( df[ order( df$cre.pos.pvalue, decreasing = F ), ] )
  df.subset <- mutate(volcano.input, sig=ifelse(volcano.input$cre.pos.pvalue < p.value.threshold, "Sig", "Not Sig"))
  
  df.subset <- df.subset [ !(is.na(df.subset$cre.pos.pvalue)), ]
  df.subset <- df.subset [ !(is.na(df.subset$cre.pos.log2fc)), ]
  df.subset <- df.subset [ df.subset$cre.pos.log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$cre.pos.log2fc != "-Inf", ]
  p.pos <- ggplot(data=df.subset, aes(x=cre.pos.log2fc, y=-log10(cre.pos.pvalue), colour= sig)) + 
  geom_point(alpha=1, size=4) +
  theme_minimal() +
  scale_color_manual( values = c( "Sig"='red3', "Not Sig" ='black' ) ) +
  theme(legend.position="none") +
  xlab("\nlog2(TAM/noTAM)") + ylab("-log10(p)\n") +
  xlim(-10,10) +
  ggtitle(paste0( title, " Cre+ (red = pvalue < 0.05)" )) +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( title, "_cre.pos.volcano.plot.png"), 1000, 800)
  print(p.pos)
  dev.off()
  
  volcano.input <- as.data.frame( df[ order( df$cre.neg.pvalue, decreasing = F ), ] )
  df.subset <- mutate(volcano.input, sig=ifelse(volcano.input$cre.neg.pvalue < p.value.threshold, "Sig", "Not Sig"))
  
  df.subset <- df.subset [ !(is.na(df.subset$cre.neg.pvalue)), ]
  df.subset <- df.subset [ !(is.na(df.subset$cre.neg.log2fc)), ]
  df.subset <- df.subset [ df.subset$cre.neg.log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$cre.neg.log2fc != "-Inf", ]
  p.neg <- ggplot(data=df.subset, aes(x=cre.neg.log2fc, y=-log10(cre.neg.pvalue), colour= sig)) + 
  geom_point(alpha=1, size=4) +
  theme_minimal() +
  scale_color_manual( values = c( "Sig"='red3', "Not Sig" ='black' ) ) +
  theme(legend.position="none") +
  xlab("\nlog2(TAM/noTAM)") + ylab("-log10(p)\n") +
  xlim(-10,10) +
  ggtitle(paste0(title, " Cre- (red = pvalue < 0.05)")) +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( title, "_cre.neg.volcano.plot.png"), 1000, 800)
  print(p.neg)
  dev.off()
  
}

volcano.plot.lsk <- function(df, title, p.value.threshold) {
  
  volcano.input <- as.data.frame( df[ order( df$pvalue, decreasing = F ), ] )
  df.subset <- mutate(volcano.input, sig=ifelse(volcano.input$pvalue < p.value.threshold, "Sig", "Not Sig"))
  
  df.subset <- df.subset [ !(is.na(df.subset$pvalue)), ]
  df.subset <- df.subset [ !(is.na(df.subset$log2fc)), ]
  df.subset <- df.subset [ df.subset$log2fc != "Inf", ]
  df.subset <- df.subset [ df.subset$log2fc != "-Inf", ]
  p <- ggplot(data=df.subset, aes(x=log2fc, y=-log10(pvalue), colour= sig)) + 
  geom_point(alpha=1, size=4) +
  theme_minimal() +
  scale_color_manual( values = c( "Sig"='red3', "Not Sig" ='black' ) ) +
  theme(legend.position="none") +
  xlab("\nlog2(Cre+/Cre-)") + ylab("-log10(p)\n") +
  xlim(-10,10) +
  ggtitle(paste0(title, "(red = pvalue < 0.05)")) +
  theme(plot.title = element_text(size=40)) +
  theme(axis.text=element_text(size=30, color="black"), axis.title=element_text(size=30, color="black")) +
  theme(axis.line = element_line(colour = "black", size=1))
  
  png(paste0( title, "_volcano.plot.png"), 1000, 800)
  print(p)
  dev.off()
  
}

volcano.plot.lsc(lsc.intensity, "lsc.intensity", 0.05)
volcano.plot.lsc(lsc.normal.intensity, "lsc.normal.intensity", 0.05)
volcano.plot.lsc(lsc.lfq, "lsc.lfq", 0.05)

volcano.plot.lsk(lsk.intensity, "lsk.intensity", 0.05)
volcano.plot.lsk(lsk.normal.intensity, "lsk.normal.intensity", 0.05)
volcano.plot.lsk(lsk.lfq, "lsk.lfq", 0.05)
```














